\HeaderA{skewhypMean}{Moments and Mode of the Skewed Hyperbolic Student's t-Distribution.}{skewhypMean}
\aliasA{skewhypKurt}{skewhypMean}{skewhypKurt}
\aliasA{skewhypMode}{skewhypMean}{skewhypMode}
\aliasA{skewhypSkew}{skewhypMean}{skewhypSkew}
\aliasA{skewhypVar}{skewhypMean}{skewhypVar}
\keyword{distribution}{skewhypMean}
\begin{Description}\relax
Functions to calculate the mean, variance, skewness and kurtosis of a
specified skewed hyperbolic t-distribution.
\end{Description}
\begin{Usage}
\begin{verbatim}
skewhypMean(mu = 0, delta = 1, beta = 1, nu = 1, param = c(mu, delta,
  beta, nu))
skewhypVar(mu = 0, delta = 1, beta = 1, nu = 1, param = c(mu, delta,
  beta, nu))
skewhypSkew(mu = 0, delta = 1, beta = 1, nu = 1, param = c(mu, delta,
  beta, nu))
skewhypKurt(mu = 0, delta = 1, beta = 1, nu = 1, param = c(mu, delta,
  beta, nu))
skewhypMode(mu = 0, delta = 1, beta = 1, nu = 1, param = c(mu, delta,
  beta, nu), tolerance = .Machine$double.eps ^ 0.5)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{mu}] Location parameter \eqn{\mu}{mu}, default is 0.
\item[\code{delta}] Peakedness parameter \eqn{\sigma}{sigma}, default is 1.
\item[\code{beta}] Skewness parameter \eqn{\beta}{beta}, default is 1. Negative
values give a negative skew, positive values give a positive skew.
\item[\code{nu}] Shape parameter \eqn{\nu}{\nu}, default is 1.
\item[\code{param}] Specifying the parameters as a vector of the form
\code{c(mu, delta, beta, nu)}.
\item[\code{tolerance}] A difference smaller than this value is taken to be zero.
\end{ldescription}
\end{Arguments}
\begin{Details}\relax
Users may either specify the values of the parameters individually or
as a vector. If both forms are specified, then the values specified by
the vector \code{param} will overwrite the other ones.

The moments are calculated as per formulae in Aas\&Haff(2006) and the
mode is calculated by numerical optimisation of the density function
using \code{\LinkA{optim}{optim}}.

Note that the mean does not exist when \eqn{\nu = 2}{nu = 2}, the
variance does not exist for \eqn{\nu \le 4}{nu <= 4}, the skewness does
not exist for \eqn{\nu \le 6}{nu <= 6}, and the kurtosis does not exist
for \eqn{\nu \le 8}{nu <= 8}.
\end{Details}
\begin{Value}
\code{skewhypMean} gives the mean of the skewed hyperbolic
t-distribution, \code{skewhypVar} the variance, \code{skewhypSkew} the
skewness, \code{skewhypKurt} the kurtosis and \code{skewhypMode} the
mode.
\end{Value}
\begin{Author}\relax
David Scott \email{d.scott@auckland.ac.nz}, Fiona Grimson
\end{Author}
\begin{References}\relax
Aas, K. and Haff, I. H. (2006).
The Generalised Hyperbolic Skew Student's t-Distribution.
In \emph{Jorunal of Financial Econometrics}, 26-Jan-2006
\end{References}
\begin{SeeAlso}\relax
\code{\LinkA{dskewhyp}{dskewhyp}}, \code{\LinkA{optim}{optim}}
\end{SeeAlso}

