---
title: "Sobol sequence implementation"
author: "CD"
date: "08/05/2022"
preamble: 
  includes: 
    \usepackage{multirow}
    \usepackage{multicol} 
    \usepackage{tabls} 
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randtoolbox)
```

Source : p311 (pdf) of Monte Carlo Methods in Financial Engineering, Glasserman (2003)

In contrast to Halton or Faure sequence, Sobol are $(t,d)$-sequences
in base 2 for all $d$ with $t$ values depending on $d$:
they are permutation of Van den Corput sequence in base 2 only.

# Principle 
$V$ a generator matrix of direction (binary)
constructed as binary expansion of numbers $v_1,\dots,v_r$.
$y$ is the state variable and $x$ its output in $[0,1]$.

Prop : $V$ is upper triangular.

For a number $k$, the binary coefficients are denoted by 
$\boldsymbol{a}(k)=(a_0(k),\dots, a_{r-1}(k))$ such that
$$k = a_0(k)2^0+\dots+ a_{r-1}(k)2^{k-1}.$$

Recurrence is 
\begin{equation}
\begin{pmatrix}y_1(k)\\ \vdots \\ y_r(k)\end{pmatrix}
=
V 
\begin{pmatrix}a_0(k)\\ \vdots \\ a_{r-1}(k)\end{pmatrix}
\text{mod } 2,
\label{eq:rec:sobol:matrix}
\end{equation}
$$
x_k = y_1(k)/2+ \dots + y_r(k)/(2^r).
$$
Special case : $V$ being the identity matrix corresponds to Van der Corput sequence.
\eqref{eq:rec:sobol:matrix} rewrites as
\begin{equation}
\boldsymbol{y}(k)
= a_0(k) v_1 \oplus a_1(k) v_2 \oplus\dots \oplus a_{r-1}(k) v_r \oplus,
\label{eq:rec:sobol:binary:ak}
\end{equation}
where $\oplus$ denotes the XOR operation.

For a $d$-dimensional sequence, we need $d$ direction numbers $(v_j)_j$.
Sobol's method relies on the use primitive polynomial of binary
coefficients 
\begin{equation}
P(x) = x^q + c_1 x^{q-1}+\dots+c_{q-1} x + 1
\label{eq:primitive:polynom}
\end{equation}
which are irreducible and the smallest power dividing the polynomial $P$
is $p=2^q-1$. the coefficients $c_q,c_0$ always equals 1.

## Examples of direction number and primitive polynomials

Polynomials up to degree 5 are given in the following table
table 5.2 of Glasserman (2003).

\begin{table}[htb!]
\centering
\begin{tabular}{clrl}
Degree & Primitive polynomial & Binary coefficients & Associated number \\
$p$ & $P(x)$ & $(c_q,\dots,c_0)=\boldsymbol{c}(k)$
& $c_q2^q+\dots+c_02^0=k$ \\
\hline
0 & 1 & 1 & 1 \\
\hline
1 & $x+1$ & $(1,1)$ & $3=2^1+2^0$\\
\hline
2 & $x^2+x+1$ & $(1,1,1)$ & $7=2^2+2^1+2^0$\\  
\hline
3 & $x^3+x+1$ & $(1,0,1,1)$ & $11=2^3+2^1+2^0$\\  
3 & $x^3+x^2+x+1$ & $(1,1,0,1)$ & $13=2^3+2^2+2^0$\\  
\hline
4 & $x^4+x+1$ & $(1,0,0,1,1)$ & $19=2^4+2^1+2^0$\\  
4 & $x^4+x^3+x^2+x+1$ & $(1,1,1,1,1)$ & $25=2^4+2^3+2^2+2^1+2^0$\\  
\hline
5 & $x^5+x+1$ & $(1,0,0,0,1,1)$ & $35=2^5+2^1+2^0$\\  
5 & $x^5+x^4+x^2+x+1$ & $(1,1,0,1,1,1)$ & $59=2^5+2^4+2^2+2^1+2^0$\\  
5 & $x^5+x^3+x^2+x+1$ & $(1,0,1,1,1,1)$ & $47=2^5+2^3+2^2+2^1+2^0$\\  
5 & $x^5+x^4+x^3+x^2+1$ & $(1,1,1,1,0,1)$ & $61=2^5+2^4+2^3+2^2+2^0$\\  
5 & $x^5+x^4+x^2+x+1$ & $(1,1,0,1,1,1)$ & $55=2^5+2^4+2^2+2+2^0$\\  
5 & $x^5+x^3+x+1$ & $(1,0,1,0,0,1)$ & $41=2^5+2^3+2^0$\\  
\hline
\end{tabular}
\caption{Primitive polynomial}
\label{tab:primitive:poly}
\end{table}

Polynomial \eqref{eq:primitive:polynom} defines a recurrence relation
between $m_j$ as
\begin{equation}
m_j = 2c_1 m_{j-1} \oplus
2^2c_2 m_{j-2} \oplus
\dots \oplus
2^{q-1}c_{q-1} m_{j-q+1} \oplus 
2^{q} m_{j-q} \oplus 
m_{j-q}.
\label{eq:rec:mj}
\end{equation}
The direction numbers are $v_j=m_j/2^j$
using \eqref{eq:rec:mj} and a set of initial values 
$m_1,\dots, m_q$.



## Examples 
Consider $k=13=2^3+2^2+2^0$ so that the primitive polynomial
is $x^3+x^2+1$.
\eqref{eq:rec:mj} writes as
$$
m_j=2m_{j-1}\oplus 8m_{j-3} \oplus m_{j-3}.
$$
Initializing with $m_1=1$, $m_2=m_3=3$ leads to 
$$
m_4 = (2\times 3)\oplus (8\times 1) \oplus 1=(1111)_2 = 15.
$$
$$
m_5 = (2\times 15)\oplus (8\times 3) \oplus 3=(00101)_2 = 5.
$$
R functions `sobol.directions.mj()` and 
`sobol.directions.vj()` compute integers $m_j$ and direction
numbers $v_j$.

```{r}
p13 <- int2bit(13)
m1<-1; m2<-m3<-3
#mj
sobol.directions.mj(c(m1,m2,m3), p13, 2, echo=FALSE, input="real", output="real")
#V matrix
head(sobol.directions.vj(c(m1,m2,m3), p13, 2, echo=FALSE, output="binary"))
```

In this example, the generator is 
$$
\boldsymbol{V}
=\begin{pmatrix}
1 & 1 & 0 & 1 & 0 \\
0 & 1 & 1 & 1 & 0 \\
0 & 0 & 1 & 1 & 1 \\
0 & 0 & 0 & 1 & 1 \\
0 & 0 & 0 & 0 & 1 \\
\end{pmatrix}
$$
This gives the following sequence for integers $k=1,2,3,31$.
$V$ produces a permutation of Van der Corput as it
uses a primitive polynomial.

\begin{tabular}{lcccc}
$k$ & $\boldsymbol{a}(k)$ & $\boldsymbol{V}\boldsymbol{a}(k)=\boldsymbol{y}(k)$ & $x_k$ \\ 
\hline
1 
&
$\begin{pmatrix}
1 \\
0 \\
0 \\
0 \\
0 \\
\end{pmatrix}$
&
$\begin{pmatrix}
1 \\
0 \\
0 \\
0 \\
0 \\
\end{pmatrix}$
& 1/2
\\
\hline
2 
&
$\begin{pmatrix}
0 \\
1 \\
0 \\
0 \\
0 \\
\end{pmatrix}$
&
$\begin{pmatrix}
1 \\
1 \\
0 \\
0 \\
0 \\
\end{pmatrix}$
& 3/4
\\
\hline
3 
&
$\begin{pmatrix}
1 \\
1 \\
0 \\
0 \\
0 \\
\end{pmatrix}$
&
$\begin{pmatrix}
0 \\
1 \\
0 \\
0 \\
0 \\
\end{pmatrix}$
& 1/4
\\
\hline
31 
&
$\begin{pmatrix}
1 \\
1 \\
1 \\
1 \\
1 \\
\end{pmatrix}$
&
$\begin{pmatrix}
1 \\
1 \\
1 \\
1 \\
1 \\
\end{pmatrix}$
& 31/32
\\
\hline
\end{tabular}

# A faster implementation using Gray code

Antanov and Saleev point out Sobol's method simplifies we use 
the Gray code representation of $k$ rather 
than the binary representation.
The Gray code is such that exactly one bit changes from $k$
to $k+1$ and is defined as 
$$
\boldsymbol{g}(k)
= \boldsymbol{a}(k) \oplus \boldsymbol{a}(\lfloor k/2\rfloor).
$$
This is a shift in the string representation.
For instance, the Gray code of 3 and 4 are
$$
\boldsymbol{g}(3)
=\boldsymbol{a}(3) \oplus \boldsymbol{a}(1)
=(011)_2\oplus(001)_2
=(010)_2,
$$
$$
\boldsymbol{g}(4)
=\boldsymbol{a}(4) \oplus \boldsymbol{a}(2)
=(100)_2\oplus(010)_2
=(110)_2.
$$
The Gray code representations of integers
$0, \dots,2^r-1$ are a permutation of the sequence of 
strings formed by the usual binary representations.
So the asymptotic property of the Gray code
$\boldsymbol{g}(k)$ remains
the same as the one of the binary representation
$\boldsymbol{a}(k)$.




\begin{tabular}{lccccccc}
 & \multicolumn{7}{c}{Binary/Gray representation} \\
integer $k$ & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
\hline
binary $\boldsymbol{a}(k)$ & 001 & 010 & 011 & 100 & 101 & 110 & 111 \\
\hline
Gray $\boldsymbol{g}(k)$ & 001 & 011 & 010 & 110 & 111 & 101 & 100 \\
\hline
\end{tabular}

Recurrence \eqref{eq:rec:sobol:binary:ak}
\begin{equation}
\boldsymbol{x}(k)
= g_0(k) v_1 \oplus g_1(k) v_2 \oplus\dots \oplus g_{r-1}(k) v_r
\label{eq:sobol:gray:gk}
\end{equation}
Consider two consecutive integers $k$ and $k+1$ differing
in the $l$th bit. So
\begin{eqnarray}
\boldsymbol{x}(k+1) 
&=&
g_0(k+1) v_1 \oplus g_1(k+1) v_2 \oplus\dots \oplus g_{r-1}(k+1) v_r \\
&=& 
g_0(k) v_1 \oplus g_1(k) v_2 \oplus\dots (g_l(k)\oplus1)v_l\dots \oplus g_{r-1}(k+1) v_r = 
\boldsymbol{x}(k+1) \oplus v_l
\label{eq:rec:sobol:gray:gk}
\end{eqnarray}
Starting from 0, we never to calculate a Gray code, only
$l$ is needed to use \eqref{eq:rec:sobol:gray:gk}.
Otherwise we need the Gray code of the starting point.

# Multivariate sequence.

Sobol initiates multivariate sequence based on hypercube distribution.
For each coordinate $i\in{1,\dots,d}$, we need a generator 
$$
\boldsymbol{V}^{(i)} = (v_1^{(i)}, \dots, v_r^{(i)}).
$$
The determinant should non zero modulo 2.

## Example in dimension 3
Consider direction numbers
$$
(m_1, m_2, m_3)
=(1,1,1),
(m_1, m_2, m_3)
=(1,3,5),
(m_1, m_2, m_3)
=(1,1,7).
$$
The associated generator matrices using the first three
primitive polynomials of Table \ref{tab:primitive:poly} are
$$
\boldsymbol{V}^{(1)}
=\begin{pmatrix} 
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
\end{pmatrix},
\boldsymbol{V}^{(2)}
=\begin{pmatrix} 
1 & 1 & 1 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
\end{pmatrix},
\boldsymbol{V}^{(3)}
=\begin{pmatrix} 
1 & 0 & 1 \\
0 & 1 & 1 \\
0 & 0 & 1 \\
\end{pmatrix}.
$$
```{r, echo=FALSE, eval=FALSE}
p13 <- int2bit(13)
#V matrix
head(sobol.directions.vj(c(1,1,7), p13, 0, echo=FALSE, output="binary"))
```

Table \ref{tab:init:mj} gives the initial values to consider up to dimension 10.



\begin{table}[htb!]
\centering

\begin{tabular}{lccccc}
$i$ & $m_1$& $m_2$& $m_3$& $m_4$& $m_5$\\
\hline
1 & 1  \\
2 & 1  \\
3 & 1 & 1 \\
4 & 1 & 3 & 7 \\
5 & 1 & 1 & 5 \\
6 & 1 & 3 & 1 & 1 \\
7 & 1 & 1 & 3 & 7 \\
8 & 1 & 3 & 3 & 9 & 9 \\
9 & 1 & 3 & 7 & 13 & 3 \\
10 & 1 & 1 & 5 & 11 & 27  \\
\hline
\end{tabular}

\caption{Initial values}
\label{tab:init:mj}
\end{table}


Using primitive polynomial of Table \ref{tab:primitive:poly}, 
the 5 direction numbers are computed. Below, we reproduce the first ten 
rows of Table 5.3 of Glasserman (2003).

```{r}
first_prim_poly <- c(1, 3, 7, 11, 13, 19, 25, 35, 59, 47)
initmj <- list(
 1, 
 1,
 c(1 , 1 ),
 c(1 , 3 , 7 ),
 c(1 , 1 , 5 ),
 c(1 , 3 , 1 , 1 ),
 c(1 , 1 , 3 , 7 ),
 c(1 , 3 , 3 , 9 , 9 ),
 c(1 , 3 , 7 , 13 , 3 ),
 c(1 , 1 , 5 , 11 , 27  ))
firstmj <- function(i)
  sobol.directions.mj(initmj[[i]], first_prim_poly[i], 
                    8-length(initmj[[i]]), echo=FALSE, 
                    input="real", output="real")

t(sapply(1:length(first_prim_poly), firstmj))
```

Page 319